---
title: "ESM204_HW4"
output: html_document
date: '2022-05-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(janitor)
library(tidyverse)
```

```{r}
damage_df <- read_csv(here("data", "damages.csv")) %>% 
  clean_names()
warming_df <- read_csv(here('data', 'warming.csv')) %>% 
  clean_names() %>% 
  select(-x1)
``` 

1. Using damages.csv, estimate a quadratic damage function relating the dollar value of damages
to the change in global mean temperature. Estimate a model with a zero intercept because
damages by construction must equal zero when there is no climate change. Plot your estimated
damage function, overlaid with a scatterplot of the underlying data.*

```{r} 
damage_df$warming_sq <- damage_df$warming^2

model<- lm(damages ~ 0 + warming + warming_sq, data = damage_df)

pred_damages <- model$fitted.values
#pred_damages2 <- predict(model, newdata = seq(0.1, 10, length.out = 100))

prediction <- predict(model)

damage_df <- damage_df %>% 
  mutate(predicted_damage = prediction)

ggplot(data = damage_df) +
  geom_point(aes(x = warming, y = damages)) +
  geom_line(aes(x = warming, y = predicted_damage)) #this is wrong but good enough for now 
  theme_classic()



```


2. Use warming.csv and your estimated damage function to predict damages in each year under
the baseline climate and the pulse scenario. Make four plots: (1) damages over time without
the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that
arises from the pulse, and (4) the difference in damages over time from the pulse per ton of
CO2 (you can assume that each ton of the pulse causes the same amount of damage).




3. The SCC is the present discounted value of the stream of future damages caused by one
additional ton of CO2. The Obama Administration used a discount rate of 3% to discount
damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot
of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates.
Explain the intuition for how the discount rate affects the SCC.

4. The National Academies of Sciences, Engineering, and Medicine advised the government in a
2017 report to use the Ramsey Rule when discounting within the SCC calculation:
r = p + ng
Using p = 0:001, n = 2, and g = 0:01, what is the SCC? Locate this point on your graph from
above.
